---
title: "SDS315_HW5"
author: "Aaron Coronado"
date: "2024-02-29"
output: html_document
---

# **Homework 5**

[GitHub Repo Link](https://github.com/aaroncoronado/SDS315_HW5.git)

```{r Setup, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)
library(kableExtra)
let_freq <- read_csv("letter_frequencies.csv")
brown <- readLines("brown_sentences.txt")
```

## **Problem 1**

$H_0$: The rate at which securities trades from the Iron Bank are flagged by the SEC's detection algorithm is equal to the baseline rate of 2.4% for all traders.\
Test Statistic: Of the last 2021 trades by Iron Bank employees, 70 were flagged by the SEC's detection algorithm.

```{r Problem 1, message = FALSE, warning = FALSE, echo = FALSE}
sim_insider = do(100000) * nflip(n = 2021, prob = 0.024)
ggplot(sim_insider) +
  geom_histogram(aes(x = nflip)) +
  labs(title = "Distribution of Simulated Flagged Trades", x = "Number of Security Tradings Flagged", y = "Frequency")
```

p-value: `r sum(sim_insider >= 70) / 100000`\
Since p-value < $\alpha$, the null hypothesis must be rejected. Therefore, the SEC should investigate the Iron Bank because 70 flagged trades out of 2021 is not plausibly consistent with the 2.4% random variability in trading patterns.

## **Problem 2**

$H_0$: The rate of health code violations at Gourmet Bites is equal to the citywide average rate of 3%.\
Test Statistic: Of the 50 inspections conducted at various branches of Gourmet Bites, 8 resulted in health code violations being reported.

```{r Problem 2, message = FALSE, warning = FALSE, echo = FALSE}
sim_violations = do(100000) * nflip(n = 50, prob = 0.03)
ggplot(sim_violations) +
  geom_histogram(aes(x = nflip), binwidth = 1) +
  labs(title = "Distribution of Simulated Violation Rates at Gourmet Bites",
       x = "Number of Violations", y = "Frequency")
```

p-value: `r sum(sim_violations >= 8) / 100000`\
Since p-value < $\alpha$, the null hypothesis must be rejected. Therefore, the Health Department should take action because Gourmet Bites' rate of health code violations is not plausibly consistent with the average 3% of random health code violations at well-managed establishments.

## **Problem 3**

### **Part A**

```{r Problem 3.A, message = FALSE, warning = FALSE, echo = FALSE}
calculate_chi_squared = function(sentence, freq_table) {
  clean_sentence = toupper(gsub("[^A-Za-z]", "", sentence))
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  return(chi_squared_stat)
}

brown_chi_sq = sapply(brown, calculate_chi_squared, freq_table = let_freq)
null_distribution = brown_chi_sq %>%
  as.tibble()

ggplot(null_distribution, aes(x = value)) +
  geom_histogram() +
  labs(title = "Distribution of Chi-Squared Values for the Brown Sentences", x = "Chi-Squared Values", y = "Frequency")
```

### **Part B**

```{r Problem 3.B, message = FALSE, warning = FALSE, echo = FALSE}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

chi_sq_vals <- sapply(sentences, calculate_chi_squared, freq_table = let_freq)

pval_table = tibble(Sentence = character(), pvalue = numeric())

for(x in 1:10){
  temp_chi_sq = chi_sq_vals[[x]]
  if(temp_chi_sq >= mean(null_distribution$value))
    temp_pval = sum(null_distribution >= chi_sq_vals[[x]]) / nrow(null_distribution)
  else
    temp_pval = sum(null_distribution < chi_sq_vals[[x]]) / nrow(null_distribution)
  pval_table <- bind_rows(pval_table, tibble(Sentence = sentences[x], pvalue = round(temp_pval, 3)))
}

kable_styling(kbl(pval_table))
```

According to my table, there are two sentences with a p value of less than 0.05. Since its p value is the lesser of the two, I think the "Feeling vexed after..." sentence was produced and watermarked by an LLM.
